
### Recode form label names

Before we begin, we shorten the form topic names by *recoding* them because they
appear to be long and disarray the plots. The *recode table* below presents a
lookup table for the form labels.

```{r recode-table}
recode_tbl %>% 
  group_by(folderName) %>% 
  gt() %>% 
  gt_condensed_style() %>% 
  tab_footnote(
    footnote = md("*Some short form names (e.g. 'Salud') stay as such, no need to make them even shorter*"),
    locations = cells_column_labels(
      columns = vars(formNameRecode)
    )
  )
```

## 2.1. Partners

They are two types of partners in the database:

+ Reporting partners: Higher level of partners reporting directly in ActivityInfo.

+ Implementing partners: Partners reporting through a reporting partner.

```{r, include=FALSE}
ppartners <- partners %>%
  group_by(partnerName) %>%
  count(partnerName, name = "freq", sort = TRUE) %>%
  ungroup() %>%
  mutate(prop = round(freq / sum(freq), 3))
```

The table below shows the count of reporting partner per each record:

+ **`r ppartners[1,"partnerName",drop=T]`** has `r xfun::n2w(ppartners[1,"freq",drop=T])`
records, 
which is `r scales::percent(ppartners[1,"prop",drop=T])`
of the total records.

+ Second, 
**`r ppartners[2,"partnerName",drop=T]`** has `r xfun::n2w(ppartners[2,"freq",drop=T])`
records, 
which is `r scales::percent(ppartners[2,"prop",drop=T])`
of the total records.

+ The most difference between percentages of the partners 
**`r ppartners[1,"partnerName",drop=T]`**
and
**`r ppartners[2,"partnerName",drop=T]`**
is
*`r paste0(floor(convolve((ppartners[1,"prop",drop=T]-ppartners[2,"prop",drop=T]), 100)), '%')`*.

<hr>

```{r}
ppartners %>%
  rename(
    `Reporting partner` = partnerName,
    `Frequency` = freq,
    `Relative frequency` = prop
  ) %>% 
  gt() %>% 
  gt_condensed_style()
```

```{r, include=FALSE}
PartnerSubPropFreq <- partners %>% 
  group_by(partnerName, subPartnerName) %>% 
  count(partnerName, subPartnerName, name = "freq", sort = TRUE) %>% 
  summarize(freq = sum(freq)) %>% 
  mutate(prop = round(freq/sum(freq), 3)) %>% 
  mutate(percent = paste0(floor(100 * freq/sum(freq)), "%"))
```

```{r, include=FALSE}
acnur.hias.part <- PartnerSubPropFreq %>% 
  dplyr::filter(partnerName %in% "ACNUR", subPartnerName %in% "HIAS")
```

<hr>

The table below shows the proportion of records entered by
partners and sub-partners.

+ *`r acnur.hias.part[,"freq",drop=T]`* out of 
*`r sum(PartnerSubPropFreq %>% dplyr::filter(partnerName %in% "ACNUR") %>% pull(freq))`* 
total responses of **ACNUR** is actually coming from **HIAS**.

+ **UNICEF** has more diversed partners in terms of reporting.
*`r PartnerSubPropFreq%>%dplyr::filter(partnerName%in% "UNICEF", subPartnerName %in% "HIAS")%>%pull(percent)`*
of responses of **UNICEF** comes from **HIAS**.
*`r PartnerSubPropFreq%>%dplyr::filter(partnerName %in% "UNICEF", subPartnerName %in% "UNICEF")%>% pull(percent)`*
of reporting comes from the **UNICEF** itself.

+ Under **PMA**, there are *`r nrow(PartnerSubPropFreq%>%dplyr::filter(partnerName %in% "PMA"))`* sub-partners. **HIAS** reports *`r PartnerSubPropFreq%>%dplyr::filter(partnerName %in% "PMA", subPartnerName %in% "HIAS")%>%pull("percent")`* of these records.

Those are the total numbers of reporting in all database, the numbers are not
specific to the narratives (multi-line text fields). In the next section, we
count the number of reportings done only in the narrative sections.

<hr>

```{r, echo=FALSE, warning=FALSE}
## Used parse & eval because of this issue: 
## https://github.com/rstudio/gt/issues/308
pspf.bold_groups <- glue::glue_collapse(glue::glue('
  tab_style(
    locations = cells_group(groups = "{x}"),
    style = cells_styles(text_weight = "bold")
  )
  ',
  x = unique(PartnerSubPropFreq$partnerName)
  ),
  sep = " %>%\n")
pspf.tbl <- PartnerSubPropFreq %>% 
  group_by(subPartnerName) %>% 
  arrange(desc(freq)) %>% 
  group_by(partnerName) %>% 
  rename(
    `Reporting partner` = partnerName,
    `Implementing partner` = subPartnerName,
    `Frequency` = freq,
    `Relative frequency` = prop
  ) %>% 
  select(-percent)
eval(parse(text=glue::glue('
pspf.tbl %>%
  gt() %>% 
  gt_condensed_style() %>% 
  {bolds}
', bolds = pspf.bold_groups)))
```

<hr>

```{r child = 'partner-tabset-partners.Rmd', eval=T}
```

## 2.2. Location

## 2.3. Narrative data

In this section, we focus on a subset of the reports, which do particularly have
the *multi-text fields*, called "Narrative data" in *ActivityInfo* terms. Plain
saying that narrative data is multi-line text fields allowing users to enter
long texts.

Note that we also keep the narratives that are empty (which are displayed as
`NA`, *Not Available*).

### The number of narrative records in form topics

```{r}
narrative.all.records.tbl <- inner_join({
  narratives %>% 
    group_by(formNameRecode) %>% 
    summarise(filled = sum(!is.na(response)),
              missing = sum(is.na(response)))
},{
  narratives %>% 
    group_by(formNameRecode) %>% 
    summarise(total = length(response))
}, by = "formNameRecode") %>%
  group_by(formNameRecode) %>% 
  mutate(`Response rate` = round(filled/missing, 3)) %>% 
  arrange(desc(filled)) %>% 
  ungroup()

n.a.r.t <- inner_join(recode_tbl, narrative.all.records.tbl, by = "formNameRecode")

narrative.all.records.tbl %>% 
  gt() %>% 
  gt_condensed_style() %>%
  tab_spanner(
    label = "Response",
    columns = vars(filled, missing, total)
  ) %>% 
  tab_footnote(
    footnote = md("*Divide the number of `filled` responses in the number of `missing` responses*"),
    locations = cells_column_labels(
      columns = vars(`Response rate`)
    )
  )
```

<br>

The table shows the narrative fields with count of *'full*' and '*empty*' fields.

```{r, results='asis'}
ss.n.a.r.t <- n.a.r.t %>% arrange(desc(total)) %>% slice(seq(5))
for (i in seq(nrow(ss.n.a.r.t))) {
  ss <- ss.n.a.r.t[i, ]
  txt <- glue::glue(
    "The form **{formNameRecode}** (*{formName}*) has *{filled}* and *{missing}* records in a total of *{total}* records.",
    formNameRecode = ss[["formNameRecode"]],
    formName = ss[["formName"]],
    filled = ss[["filled"]],
    missing = ss[["missing"]],
    total = ss[["total"]]
  )
  cat("+", txt, "\n\n")
}
```

Please note that the following form topics are not included in the table above
because they do not contain any narrative fields:
`r paste("*",setdiff(unique(form.table$formName), unique(narratives$formName)), "*", sep = "", collapse = ", ")`

### Partners entering narrative data

The number of 'Reporting Partners' and 'Implementing Partners' reporting narrative
(multi-line text) data:

```{r}
narratives.part.count <- narratives %>% 
  group_by(partnerName, subPartnerName) %>% 
  count(partnerName, subPartnerName, name = "freq", sort=TRUE) %>% 
  summarize(freq = sum(freq)) %>% 
  mutate(prop = round(freq/sum(freq), 3)) %>% 
  mutate(percent = paste0(floor(100 * freq/sum(freq)), "%"))

## partners having multiple sub-partners
narratives.part.count.with.subPartners <- narratives.part.count %>% 
  dplyr::filter(prop != 1) %>% 
  ungroup()

## partners do not have any sub-partners reporting via them.
narratives.part.count.without.subPartners <- narratives.part.count %>% 
  dplyr::filter(prop == 1) %>% 
  ungroup()
```

As we have seen previously, Not all Reporting and Implementing Partners record
multi-line narrative textual data. For instance, the partner **PMA** has lots of
Implementing Partners reporting for the different data types (as seen in above)
but there are no narratives from them. `TODO fix NA`

```{r, warning=FALSE, fig.width=7, fig.height=8}
## grouped facet wraps. Source:
## https://stackoverflow.com/questions/14840542/place-a-legend-for-each-facet-wrap-grid-in-ggplot2

plot_partners_splitted <- split(
  x = narratives.part.count.with.subPartners, 
  f = narratives.part.count.with.subPartners$partnerName
)

## bring ACNUR and UNICEF plots to the front:
plot_partners_splitted_front <- plot_partners_splitted[c(
  c("ACNUR", "UNICEF"),
  setdiff(names(plot_partners_splitted), c("ACNUR", "UNICEF"))
)]

plot_partners_list <- lapply(seq_along(plot_partners_splitted_front), function(i) {
  ggplot(plot_partners_splitted_front[[i]], aes(x = partnerName, y = prop)) +
    geom_bar(aes(fill = reorder(subPartnerName,-prop)), position = "dodge", stat =
               "identity", color = "black") +
    facet_wrap(~ partnerName, scales = "free") +
    geom_label(
      aes(
        label = percent,
        group = reorder(subPartnerName,-prop),
        vjust = 0.65
      ),
      position = position_dodge(width = 1),
      label.size = 0.35
    ) +
    ylim(0, 1) + 
    xlab(NULL) +
    ylab(NULL) +
    theme_ecuador1(border = TRUE) +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      legend.title = element_blank(),
      legend.position = "bottom"
    ) +
    scale_fill_brewer(palette = eval(quote(brewer_diverging_color_names()[i])))
})

## "cowplot" is civilized version of 'gridExtra'.
do.call(cowplot::plot_grid, plot_partners_list)
```

+ *`r subset(narratives.part.count.with.subPartners, partnerName == "ACNUR" & subPartnerName == "HIAS", "percent", drop = T)`*
of the narrative records are
entered by the implementing partner **HIAS** reported via **ACNUR**.
Only
*`r subset(narratives.part.count.with.subPartners, partnerName == "ACNUR" & subPartnerName == "ACNUR", "percent", drop = T)`*
of the narrative records are entered by **ACNUR** itself.

+ Also **HIAS** enters *`r subset(narratives.part.count.with.subPartners, partnerName == "UNICEF" & subPartnerName == "HIAS", "percent", drop = T)`* of the narrative records via **UNICEF**.

+ The rest of the reporting partners do not have implementing partners, they do
the all implementation themselves:
*`r paste(unique(narratives.part.count.without.subPartners$partnerName), collapse = ", ")`*.

### The number of cantons and provinces entering narrative data

```{r, include = FALSE}
locations <- narratives %>% 
  count(province, canton, name = "freq") %>%
  group_by(province, canton) %>% 
  summarize(freq = sum(freq)) %>%
  mutate(canton.prop = round(freq/sum(freq), 3)) %>% 
  mutate(canton.percent = paste0(floor(100 * freq/sum(freq)), "%")) %>% 
  ungroup() %>% 
  mutate(province.prop = round(freq/sum(freq), 3)) %>% 
  mutate(province.percent = paste0(floor(100 * freq/sum(freq)), "%")) %>% 
  arrange(desc(freq))
```

```{r canton-locations-tbl}
## Used parse & eval because of this issue: 
## https://github.com/rstudio/gt/issues/308
lcts.bold_groups <- glue::glue_collapse(glue::glue('
  tab_style(
    locations = cells_group(groups = "{x}"),
    style = cells_styles(text_weight = "bold")
  )
  ', 
  x = unique(locations$province)
),
sep = " %>%\n")
lcts.tbl <- locations %>%
  select(province, canton, freq, canton.prop, province.prop) %>%
  group_by(province) %>% 
  arrange(desc(freq)) %>% 
  rename(
    `Canton` = canton,
    `Frequency` = freq,
    `Relative frequency (canton)` = canton.prop,
    `Relative frequency (province)` = province.prop
  )
eval(parse(text=glue::glue('
lcts.tbl %>% 
  gt() %>% 
  tab_header(
    title = "Canton and provinces",
    subtitle = "The number of reports in the multi-text (narrative) fields"
  ) %>% 
  gt_condensed_style() %>% 
  {bolds}
', bolds = lcts.bold_groups)))
```

<br>

```{r child = 'treemap-canton.Rmd', eval=T}
```
