
## The nature of data

This section gives some ideas how the raw data of Ecuador monitoring looks like. 

* The data has been
extracted from *ActivityInfo* by using the *ActivityInfo API* and pre-processed
to make it ready for the analysis. 

* The most of data extraction and cleaning are
done beforehand (please see `R/` folder in the repository especially take a
close look at `etl.R` and `etl-methods.R` files). If you want to download the
raw data, you must have an access for it, that can be done by sourcing the
`etl.R` file.

* Some description about the nature of data:
  
  + `databaseId`: the internal ActivityInfo id for databases
    
  + `databaseName`: the name of databases visible to users
  
  + `folderId`: the internal ActivityInfo id for folders
  
  + `folderName`: the name of folders visible to users
  
  + `formId`: the internal ActivityInfo id for forms
  
  + `formName`: the name of forms visible to users
  
  + `subFormId`: the internal ActivityInfo id for the sub-forms where the records are kept
  
  + `subFormName`: the name of the sub-forms visible to users
  
  + `Month`: indicating month when a record is entered
  
  + `code`: Schema question code
  
  + `question`: Question label indicated by the code
  
  + `response`: Response given by users
  
  + `required`: A boolean value to check whether the question is required to complete.
  
  + `type`: internal type for the code. The available types in the data are 
  `r unique(form.table$type)`.
  
  + `partnerName`: The name of reporting partners. The name of implementing
  partners can be extracted from the data.
  
  + `canton`: The canton name of the record.
  
  + `province`: The province name of the record.
  
  + `description`: the description field further explaining what the question mean.
  The cells are represented as `NA` when fields not exists or not applicable.

<hr>

## Technical aspects of the data

```{r, include=FALSE}
library(inspectdf)
```

For this analysis, we use the
[**inspectdf**](https://cran.r-project.org/package=inspectdf) package.

```{r}
inspect_types(form.table) %>%
  show_plot()
```

+ Almost all columns are character vector, except the column
*`r colnames(form.table)[sapply(form.table, class) %in% "logical"]`*.

```{r}
inspect_cat(form.table) %>%
  show_plot()
```

Above, we look at the categorical features of the data frame. That shows the each
categorical column, number of unique levels, and the most common levels.

```{r}
# TODO fix
# inspect_na(form.table) %>%
#   show_plot()
```

The prevalance of NAs (missing values) in the colums.

```{r}
inspect_mem(form.table) %>%
  show_plot()
```

This part may seem too technical, but can be important for the other data
analyst that would work on this data.
As seen in the plot, the "response" column has the most size because it holds
long text fields.

<hr>

You can also click at the button to have a glimpse of the raw data.

<button class="btn btn-primary" data-toggle="collapse" data-target="#dataPeak">Details</button>
<div id="dataPeak" class="collapse">  

```{r, paged.print=FALSE}
peak.ft <- as.data.frame(head(form.table, 4))
peak.ft <- rbind(peak.ft, "...")
peak.ft
```

</div>

<br>

Please see the
[**ActivityInfo documentation**](http://help.activityinfo.org/m/quick_start_database_design/c/240396)
for more information about how the information is structured.

