
# How much does the average word count differ between sectors? {-#average-word-count-per-sector}

```{r}
narrative_mean_diff <- local({
  nrform <- narratives.wc %>% 
    group_by(formNameRecode)
  median <- nrform %>% 
    summarise(mean = mean(.responseWordCount))
  sampleSize <- nrform %>% 
    tally(name = "sampleSize")
  sd <- nrform %>% 
    summarise(SD = sd(.responseWordCount, na.rm=TRUE))
    
  median %>% 
    left_join(sampleSize, by = "formNameRecode") %>% 
    left_join(sd, by = "formNameRecode")
})
```

```{r, fig.width=10,fig.height=8,fig.align='center', fig.cap='Average word count and sample size'}
narrative_mean_diff %>% 
  ggplot(aes(sampleSize, mean)) + 
  geom_point() +
  ggrepel::geom_text_repel(data = narrative_mean_diff[narrative_mean_diff$sampleSize <= 100, ], 
                           aes(label = formNameRecode), segment.colour = NA, force = .2) +
  labs(
    title = "Compare average word count and sample size between sectors",
    subtitle= "Scatter plot to compare averages in word counts with sample sizes"
  ) +
  xlab("Sample size") +
  ylab("Average word count") +
  gghighlight::gghighlight(sampleSize > 100, label_key = formNameRecode,
              unhighlighted_colour = ggplot2::alpha("darkgrey", 1)
              ) +
  theme_ecuador1()
```
  
Looking at the sample size of each sector is important to see the variation
between them because some of them just look like that they are based on very
small observations.
 
It's important of being careful before making to any conclusions on a single
graph that shows a single statistics of word count on all sectors. Here is a
plot showing the average of word counts in sectors with their sample size, which
is the number of narrative records in the group.

In that consideration, for text mining, it would be better to consider the
sectors with which the average word count value is greater than and equal to
*100*. Therefore, those sectors are:
```{r}
narrative_mean_diff_text_writer <- function(tbl, limit) {
  sub <- tbl[tbl$sampleSize >= limit,]
  txts <- sapply(seq(nrow(sub)), function(i) {
    row <- sub[i, ]
    sprintf(
      "+ **%s** (M = *%.2f*, SD = *%.2f*) with *%i* sample size",
      row[["formNameRecode"]],
      row[["mean"]],
      row[["SD"]],
      row[["sampleSize"]]
    )
  })
  paste(txts, collapse = "\n\n\n")
}
```
```{r, results='asis'}
cat(narrative_mean_diff_text_writer(narrative_mean_diff, limit = 100))
```

