
# Technical aspects of the data {-}

```{r load-inspect-df, include=FALSE}
library(inspectdf)
```

This part might seem to be too technical for readers who are mainly interested
in the results. The reason why this page exists is that the information here can
be essential for the analysts working with the data.

The [*inspectdf*](https://cran.r-project.org/package=inspectdf) package is used
to generate the visualizations in this page.

<hr>

```{r, fig.align='center', fig.cap='The column types in the data frame'}
inspect_types(form.table) %>%
  show_plot()
```

Almost all columns are character vector, except 
*`r colnames(form.table)[sapply(form.table, class) %in% "logical"]`*.

<hr>

```{r, fig.align='center', fig.cap='Summarising categorical features in the data frame'}
inspect_cat(form.table) %>%
  show_plot()
```

We see that the *response* field has many unique levels which has a solid color
in the chart.

<hr>

```{r, fig.align='center', fig.cap='See the most frequently occurring categorical level in each column.'}
inspect_imb(form.table) %>% 
  show_plot()
```

<hr>

```{r, fig.align='center', fig.cap='The prevalance of NAs (missing values) in the colums.'}
inspect_na(form.table) %>%
  show_plot()
```

<hr>

```{r, fig.align='center', fig.cap='The size of each column in the data frame.'}
inspect_mem(form.table) %>%
  show_plot()
```

As seen in the plot, the *response* column has the most size because it holds
long text fields.

<hr>

Look at below to have a glimpse of the raw data.

```{r, paged.print=FALSE}
peak.ft <- as.data.frame(head(form.table, 4))
peak.ft <- rbind(peak.ft, "...")
tibble::as_tibble(peak.ft)
```

