
# Technical aspects of the data {-}

```{r, include=FALSE}
library(inspectdf)
```

For this analysis, we use the
[**inspectdf**](https://cran.r-project.org/package=inspectdf) package.

```{r, fig.align='center'}
inspect_types(form.table) %>%
  show_plot()
```

+ Almost all columns are character vector, except the column
*`r colnames(form.table)[sapply(form.table, class) %in% "logical"]`*.

```{r, fig.align='center'}
inspect_cat(form.table) %>%
  show_plot()
```

Above, we look at the categorical features of the data frame. That shows the each
categorical column, number of unique levels, and the most common levels.

```{r, fig.align='center'}
inspect_na(form.table) %>%
  show_plot()
```

The prevalance of NAs (missing values) in the colums.

```{r, fig.align='center'}
inspect_mem(form.table) %>%
  show_plot()
```

This part may seem too technical, but can be important for the other data
analyst that would work on this data.
As seen in the plot, the "response" column has the most size because it holds
long text fields.

<hr>

Look at below to have a glimpse of the raw data.

```{r, paged.print=FALSE}
peak.ft <- as.data.frame(head(form.table, 4))
peak.ft <- rbind(peak.ft, "...")
tibble::as_tibble(peak.ft)
```

