
# Data in ActivityInfo over time {-}

```{r}
ICT <- tibble(
  Date = c("Nov 2013", "June 2015", "Sept 2016", "May 2017", "May 2019"),
  Quantity = c(141442, 1887857, 3380991, 4932977, 12174327),
  Select = c(30531, 745841, 1296548, 1809419, 7595829),
  `Single-line text` = c(NA, 85863, 191640, 265196, 2683945),
  `Multi-line text` = c(6309, 57128, 116184, 168599, 915948),
  `% of total data collected` = c("3.54%", "2.06%", "2.33%", "2.35%", "3.92%")
)

## replace NAs with 0
ICT[is.na(ICT)] <- 0

ICT %>%
  gt() %>%
  fmt_number(
    columns = vars(Quantity, Select, `Single-line text`, `Multi-line text`),
    decimals = 0,
    use_seps = T
  ) %>% 
  tab_header(
    title = md("Indicator count totals"),
    subtitle = sprintf("%s to %s", ICT$Date[1], ICT$Date[length(ICT$Date)])
  ) %>% 
  cols_align(
    align = "center",
    columns = names(ICT)
  )
```

ActivityInfo has different field types for entering data.

+ [Quantity field](http://help.activityinfo.org/m/75942/l/852515-quantity-field)

  > Quantity fields allow users to enter a numerical value. 

+ [Single selection field](http://help.activityinfo.org/m/75942/l/852519-single-selection-field)
  
  > Single selection fields can be used to ask users to select one out two or more options as an answer.

+ [Multiple selection field](http://help.activityinfo.org/m/75942/l/852998-multiple-selection-field)

  > Multiple selection fields can be used to allow users to select more than one option as an answer.

+ [Multi-Line Text Field](http://help.activityinfo.org/m/75942/l/852517-multi-line-text-field)

  > Multi-line text fields can be used to collect long answers to open-ended questions.

<br>

```{r cumulative-plot, fig.align='center', fig.cap='Total multi-line text entries collected by year (cumulative)'}
ggplot(ICT, aes(reorder(Date, `Multi-line text`), `Multi-line text`)) +
  geom_bar(stat = "identity", fill = "#3fae29") +
  xlab(NULL) +
  ylab("Number of values") +
  labs(title = "Total multi-line text entries collected by year (cumulative)") +
  scale_y_continuous(labels = scales::comma) +
  theme_ecuador1()
```

<br>

```{r data-type-plot, fig.align='center', fig.cap='Total collected data by type (cumulative)'}
ICT.wide <- ICT %>% 
  select(-`% of total data collected`) %>% 
  gather(variable, value, -Date, Quantity, Select, `Single-line text`, `Multi-line text`) %>% 
  mutate(variable = as.factor(variable))

ggplot(ICT.wide, aes(reorder(Date, value), value, fill = variable)) +
  geom_bar(stat = "identity", position = position_stack(reverse = F)) +
  xlab(NULL) +
  ylab("Number of values") +
  labs(title = "Total collected data by type (cumulative)") +
  scale_y_continuous(labels = scales::comma) +
  theme_ecuador1() +
  theme(legend.title = element_blank()) +
  scale_fill_brewer(palette = "Set1")
```

> From the perspective of *ActivityInfo*, it shows a clear need for new tools to
support analysis of qualitative data (or namely 'Multi-line text entries') as
the absolute volume of qualitative data has increased by a factor of 150, and
almost doubled as a relative share of all data collected.

