
# The Description field {-}

In *ActivityInfo* forms, the questions can have an additional description field which
is meant to give extra details about the questions.

The description field is not mandatory, so it is optional, but it may
help users filling out the form easily and eliminate potential confusions if
they are leading the users enough.

Do some questions with the extra description field have better *response
quality* than the questions which do not have it?

Looking at the table containing sector name, question and description fields and
so on others:

```{r, table1}
description.field.wc <- narratives.wc %>% 
  mutate(.hasDescription = ifelse(is.na(description), FALSE, TRUE)) %>% 
  select(formName, formNameRecode, 
         .responseWordCount, .questionWordCount, .descriptionWordCount, 
         .hasDescription)
```

```{r, fig.width=10, fig.height=6, fig.align='center'}
desc.resp.wc.plot <- description.field.wc %>% 
  ## made 'has description' column factor to modify facets.
  mutate(.hasDescription = factor(
    description.field.wc$.hasDescription, 
    levels = c(TRUE, FALSE),
    labels = c("TRUE", "FALSE"))
  )

desc.resp.wc.plot.bg <- desc.resp.wc.plot %>%
  select(formNameRecode, .responseWordCount)

desc.resp.wc.plot.single.stat <- desc.resp.wc.plot %>%
  group_by(.hasDescription) %>%
  summarise(Mean.rwc = mean(.responseWordCount)) %>% 
  arrange(.hasDescription)

desc.resp.wc.plot.colors <- c("#00BFC4", "#F8766D")

ppd <- ggplot(data = desc.resp.wc.plot,
              aes(x = .responseWordCount, y = formNameRecode, colour = .hasDescription)) +
  geom_point(data = desc.resp.wc.plot.bg,
             colour = "grey",
             alpha = 0.2) +
  geom_point(size = 3L) +
  geom_point(shape = 21, color = "black", stroke = 0.2, size = 3L) +
  geom_vline(
    data = desc.resp.wc.plot.single.stat,
    mapping = aes(xintercept = Mean.rwc, color = .hasDescription),
    linetype = "dashed"
  ) +
  facet_wrap( ~ .hasDescription) +
  scale_colour_manual(values = desc.resp.wc.plot.colors) +
  theme_ecuador1(border = TRUE) +
  guides(colour = FALSE) +
  labs(
    title = "Word count of responses depending upon the presence of description field",
    subtitle = sprintf(
      "Average response word count of questions with a description (%.2f) without any description (%.2f) field",
      dplyr::filter(desc.resp.wc.plot.single.stat, .hasDescription == TRUE)[["Mean.rwc"]],
      dplyr::filter(desc.resp.wc.plot.single.stat, .hasDescription == FALSE)[["Mean.rwc"]]
    )
  ) +
  xlab("Word count of responses") +
  ylab(NULL)

## The solution to separately color facet headings is taken from here:
## https://github.com/tidyverse/ggplot2/issues/2096#issuecomment-292971159
g <- ggplot_gtable(ggplot_build(ppd))
stripr <- grep('strip-t', g$layout$name)
k <- 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- desc.resp.wc.plot.colors[k]
  k <- k+1
}
grid::grid.draw(g)
```

<br>

We see in the plot above that the response word counts per form and colored if a
response has a description field or not. Having a description field or not is
calculated as that a description field has a minimum one word.

The responses with the longest word counts are the ones with description.
Nevertheless, it is not so easy to see a clear trend that there's a correlation
between response word count and description fields. Interestingly, the form
topic *Protección_VBG* has no description fields at all in its form topics.

