
# Distribution of word counts in response fields {-}

We will focus on word count of response field because the data we are interested
in lies there.

There's a lot of variation in response word count. The response word count has a
distribution with a tail is on the left side meaning that most of the values are
accumulated on the left. So this distribution has a negative skew.

```{r, fig.align='center', fig.width=8, fig.height=7, fig.cap='Distribution of word count of responses. A zoom to response word counts between 0 and 100.'}
narratives.wc %>% 
  ggplot(aes(.responseWordCount)) +
  geom_histogram(bins = 100, fill = "steelblue", color = "black", size = 0.1) +
  xlab("Number of words") +
  ggforce::facet_zoom(x = .responseWordCount <= 100)
```

If we look at the distribution of word count of responses by *specific sectors*
which have relatively higher sample size:

```{r, fig.align='center', fig.width=8, fig.height=5, fig.cap='Distribution of word count of responses in "VBG", "Poblacion" and "Tecnico" sectors'}
narratives.wc %>% 
  dplyr::filter(formNameRecode %in% c("VBG", "Poblacion", "Tecnico")) %>% 
  ggplot(aes(.responseWordCount)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "black", size = 0.1) +
  facet_wrap(~formNameRecode) +
  xlab(NULL) +
  ylab(NULL) +
  theme_ecuador1(border = TRUE)
```

`TODO`
