# Dates and years in the responses {-}

```{r LOCALE_SETUP, include=FALSE}
## temporarily change the system locale:
old.locale <- Sys.getlocale("LC_TIME")
Sys.setlocale("LC_TIME", "es_ES.UTF-8")
new.locale <- Sys.getlocale("LC_TIME")
stopifnot(identical(new.locale, "es_ES.UTF-8")) ## mini test
```

The dates used in the responses can refer to an important information that is
being neglected in the response fields. If certain dates are repeatedly used in
the text, it should be time to open a new field for it in order to make
monitoring process more robust and efficient.

The tool to look and extract dates in the text is *regular expressions*. The
regular expressions[^10], shortly *RegEx*, are character patterns that are used
to search and manipulate strings.

Although the most common pattern to write dates in Spanish follows the rule of
*day* - *month* - *year*, not all dates are written in a consistent pattern.

**A good rule of thumb to make this analysis easier, adapt writing dates in the
*ISO 8601 format*[^11], which is an international standard for writing dates, in
the future.**

Therefore, we set up a complicated pattern to extract the strings, which covers
the following cases below:

The explanation of sequences
(the descriptions adapted from the *POSIX standard*[^12]):

| sequence | explanation |
|:---------|:------------|
| (sign) | a sign such as `~!#$%^&*()_-+=` etc. |
| `d` | day of the month as decimal number (01–31) |
| `m` | month as decimal number (01–12) |
| `Y` | year |
| `B` | locale's full month name (e.g., January) |
| `b` | locale's abbreviated month name (e.g., Jan) |

You might try any of these sequences in R by using `format` such as 
`format(as.Date("25-09-2019"), "%B")`.

```{r}
formats <- list(
  format1 = list(
    regex = "(19|20)\\d{2}",
    description = "Most of the years start with 19xx and 20xx, so better to choose with this prefix to reduce the number of false positives which are just numbers with four digits."
  ),
  format2 = list(
    regex = "(?:enero|febrero|marzo|abril|mayo|junio|julio|agosto|septiembre|octubre|noviembre|diciembre)",
    description = "Months in Spanish"
  ),
  format3 = list(
    regex = "\\d{2}[-|/]\\d{2}[-|/]\\d{4}",
    description = "day month and year separated by a hyphen or a forward slash"
  ),
  format4 = list(
    regex = "(?:ene(?:ro)?|feb(?:rero)?|mar(?:zo)?|abr(?:il)?|may(?:o)?|jun(?:io)?|jul(?:io)?|ago(?:sto)?|sep(?:tiembre)?|oct(?:ubre)?|nov(?:iembre)?|dic(?:iembre)?)\\s+(?:19[7-9]\\d|2\\d{3})",
    description = "full or abbreviated month name and year"
  ),
  format5 = list(
    regex = "(?:ene(?:ro)?|feb(?:rero)?|mar(?:zo)?|abr(?:il)?|may(?:o)?|jun(?:io)?|jul(?:io)?|ago(?:sto)?|sep(?:tiembre)?|oct(?:ubre)?|nov(?:iembre)?|dic(?:iembre)?)\\s+\\d+.*(?:19[7-9]\\d|2\\d{3})",
    description = "full or abbreviated month name and year and day and year"
  ),
  format6 = list(
    regex = "\\d+\\s+de\\s+(?:ene(?:ro)?|feb(?:rero)?|mar(?:zo)?|abr(?:il)?|may(?:o)?|jun(?:io)?|jul(?:io)?|ago(?:sto)?|sep(?:tiembre)?|oct(?:ubre)?|nov(?:iembre)?|dic(?:iembre)?)\\s+(?:19[7-9]\\d|2\\d{3})",
    description = "A common way writing dates in Spanish"
  ),
  format7 = list(
    regex = "\\d+\\s+de\\s+(?:ene(?:ro)?|feb(?:rero)?|mar(?:zo)?|abr(?:il)?|may(?:o)?|jun(?:io)?|jul(?:io)?|ago(?:sto)?|sep(?:tiembre)?|oct(?:ubre)?|nov(?:iembre)?|dic(?:iembre)?)\\s+de\\s+(?:19[7-9]\\d|2\\d{3})",
    description = "A common way writing dates in Spanish"
  ),
  format8 = list(
    regex =  "\\d+\\s+de\\s+(?:ene(?:ro)?|feb(?:rero)?|mar(?:zo)?|abr(?:il)?|may(?:o)?|jun(?:io)?|jul(?:io)?|ago(?:sto)?|sep(?:tiembre)?|oct(?:ubre)?|nov(?:iembre)?|dic(?:iembre)?)\\s+del\\s+(?:19[7-9]\\d|2\\d{3})",
    description = "A common way writing dates in Spanish"
  )
)
```

| index | format | variants | RegEx | RegEx description |
|:------|:-------|:---------|:------|:------------------|
| 1 | `Y` | `2019` | `r formats$format1$regex` | `r formats$format1$description` |
| 2 | `B` | `septiembre` | `r formats$format2$regex` | `r formats$format2$description` |
| 3 | (`d` or `m`)(sign)(`m` or `d`)(sign)`Y` | `25/09/2019`, `25-09-2019`, `09-25-2019` etc. | `r formats$format3$regex` | `r formats$format3$description` |
| 4 | `B` `Y` | `septiembre 2019`, `sep 2019` |  `r formats$format4$regex` | `r formats$format4$description` |
| 5 | (`B` or `b`) `d`(sign) `Y` | `julio 19 2019`, `jul 9, 2019`, `jul 9. 2019` | `r formats$format5$regex` | `r formats$format5$description` |
| 6 | `d` de `B` `Y` | `28 de marzo 2019` | `r formats$format6$regex` | `r formats$format6$description` |
| 7 | `d` de `B` de `Y` | `16 de octubre de 2019` | `r formats$format7$regex` | `r formats$format7$description` |
| 8 | `d` de `B` del `Y` | `1 de octubre del 2019` | `r formats$format8$regex` | `r formats$format8$description` |

After the texts are extracted, they are displayed in a form of that they
position of dates are shown in the sentences or paragraphs together with the
adjacent words.

Spanish month names and abbreviated version in *POSIX standard*:
```{r}
all.Spanish.months <- data.frame(
  full = c("enero", "febrero", "marzo", "abril", "mayo", "junio", "julio", 
           "agosto", "septiembre", "octubre", "noviembre", "diciembre"),
  abbreviated = c("ene", "feb", "mar", "abr", "may", "jun", "jul", "ago", "sep", 
                  "oct", "nov", "dic"),
  stringsAsFactors = FALSE
)
all.Spanish.months %>% 
  gt()
```

```{r}
rps <- narratives[!is.na(narratives$response),]
seq_formats <- seq(8L)
rpss <- unique(do.call(rbind, lapply(seq_formats, function(i) {
  fname <- paste0("format", i)
  format <- rps[grep(formats[[fname]][["regex"]],
                     rps$response, 
                     perl = TRUE), 
                "response"]
})))
matches <- lapply(seq_formats, function(i) {
  fname <- paste0("format", i)
  regmatches(rps$response, gregexpr(formats[[fname]][["regex"]], rps$response, perl = TRUE))  
})
rpss.full <- rps %>% 
  inner_join(rpss, by = "response")
rpss.full %>% 
  head()
```

The found dates are as follows:
```{r}
found_dates <- do.call(c, purrr::flatten(Filter(function(x) length(x) > 0, purrr::flatten(matches))))
found_dates_unique <- unique(found_dates)
found_dates_unique %>% 
  enframe() %>% 
  select(value) %>% 
  gt()
```

The most common dates used in the responses:
```{r}
enframe(found_dates) %>% 
  count(value) %>% 
  arrange(desc(n)) %>% 
  slice(seq(20)) %>% 
  gt()
```

```{r}
rpss_words <- rpss.full %>% 
  filter(formNameRecode %in% c("VBG", "Poblacion", "Tecnico")) %>% 
  mutate(matching_word = as.character(stringr::str_match(response, paste(found_dates_unique, collapse = "|")))) %>% 
  unnest(cols = c(matching_word))
rpss_words %>% 
  group_by(formNameRecode) %>% 
  count(matching_word) %>% 
  arrange(desc(n)) %>% 
  ggplot(aes(matching_word, n, color = formNameRecode)) +
  geom_point() +
  ggplot2::coord_flip() +
  xlab(NULL) +
  ylab("Number of occurences") +
  labs(
    title = "The dates in sectors",
    subtitle = "The chosen sectors are those with high sample sizes: *VBG*, *Poblacion*, and *Tecnico*"
  ) +
  theme_ecuador1() +
  theme(legend.title = element_blank()) +
  theme(
    plot.title = ggtext::element_markdown(lineheight = 1.1),
    plot.subtitle = ggtext::element_markdown(lineheight = 1.1)
  )
```

```{r, fig.cap="Since this document is being exposed to online, full versions of texts are not displayed. The full version is available upon a valid request.", fig.width=7, fig.height=2, fig.align="center"}
library(ggpage) # fun to create with ggpage
rpss.full_ggpage <- 
  rpss.full %>% 
  select(recordId, response) %>% 
  dplyr::filter(nchar(response) < median(nchar(rpss.full$response))) %>% # filter bcs long responses breaks
  transmute(text = response, book = recordId)

rpss.full_ggpage %>% 
  ggpage_build(character_height = 7) %>% 
  mutate(matching_word = stringr::str_detect(word, paste(found_dates_unique, collapse = "|"))) %>% 
  ggpage_plot(aes(fill = matching_word)) +
  scale_fill_manual(values = c("grey80", "red"), 
                    labels = c("No date", "Date"),
                    name = "Dates in the responses") +
  theme(legend.position = "bottom")
  
```

[^10]: More precisely, the RegEx engine in R is called as *P*erl-*C*ompatible
*R*egular *E*xpressions (PCRE) <https://www.pcre.org/original/doc/html/>
Accessed September 9, 2019.

[^11]: ISO 8601 Date and Time Format
<https://www.iso.org/iso-8601-date-and-time-format.html> Accessed September 9,
2019.

[^12]: POSIX 1003.1 - man page for date (posix section 1p)
<https://www.unix.com/man-page/posix/1p/date/> Accessed September 9, 2019.

```{r LOCALE_BACK_SETUP, include=FALSE}
## set-up the changed locale back:
Sys.setlocale("LC_TIME", old.locale)
```

