# Median word count of responses {-}

Before we begin, a good note to keep in mind is that reducing multiple values
down to a single value should be avoided in the early stages of the analysis
because reducing hides a lot e.g. a bar chart showing average the word count per
partner. Some partners may write longer than others, because:

1. They *actually* write longer than other partners.

2. The questions they answered require short answers.

This is something to be known together with the *qualitative data analysis*.

<br>

#### Median word count of responses per sector name {-}

```{r, fig.align='center'}
x1<-narratives.wc %>% 
  dplyr::filter(formNameRecode != "Salud") %>% 
  group_by(formNameRecode) %>% 
  summarise(median = median(.responseWordCount)) %>% 
  ungroup()

x1 %>%   
  ggplot(aes(reorder(formNameRecode, -median), median)) +
  geom_bar(stat = "identity", width = 0.5, fill = "tomato2", color = "black") + 
  labs(
    title = "Median word count of responses per sector name",
    subtitle = sprintf("The average median rate %.2f", mean(x1$median))
  ) +
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
  xlab(NULL) +
  ylab(NULL)
```

In the plot above, we removed the sector **Salud** sector because it has only 
*`r xfun::n2w(nrow(narratives.wc[narratives.wc$formNameRecode=="Salud",]))`* 
records and skews the data.

<br>

#### Median word count of responses per Reporting Partner {-}

```{r, fig.align='center'}
narratives.wc %>% 
  group_by(partnerName) %>% 
  summarise(median = median(.responseWordCount)) %>% 
  ungroup() %>% 
  ggplot(aes(reorder(partnerName, -median), median)) +
  geom_bar(stat = "identity", width = 0.5, fill = "goldenrod", color = "black") + 
  labs(
    title = "Median word count of responses per partner",
    subtitle = sprintf("The average median rate %.2f", mean(narratives.wc$.responseWordCount))
  ) +
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
  xlab(NULL) +
  ylab(NULL)
```

<br>

#### Median word count of responses per specific Reporting Partners {-}

```{r, fig.align='center'}
narratives.wc.median.specific.partners <- narratives.wc %>% 
  group_by(partnerName, subPartnerName) %>% 
  summarise(median = median(.responseWordCount)) %>% 
  dplyr::filter(partnerName %in% c("ACNUR", "UNICEF") | subPartnerName %in% "HIAS")

narratives.wc.median.specific.partners_splitted <- split(
  x = narratives.wc.median.specific.partners, 
  f = narratives.wc.median.specific.partners$partnerName
)

narratives.wc.median.specific.partners_splitted_list <- 
  lapply(seq_along(narratives.wc.median.specific.partners_splitted), function(i) {
    ggplot(narratives.wc.median.specific.partners_splitted[[i]], aes(x = partnerName, y = median)) +
      geom_bar(aes(fill = reorder(subPartnerName,-median)), position = "dodge", stat =
                 "identity", color = "black") +
      facet_wrap(~ partnerName, scales = "free") +
      #ylim(0, 1) + 
      xlab(NULL) +
      ylab(NULL) +
      theme_ecuador1(border = TRUE) +
      theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom"
      ) +
      scale_fill_brewer(palette = eval(quote(brewer_diverging_color_names()[i])))
  })

## "cowplot" is civilized version of 'gridExtra'.
cowplot::plot_grid(plotlist = narratives.wc.median.specific.partners_splitted_list)
```

<br>

#### Median word count of responses per Reporting Partner over months {-}

```{r, include=FALSE}
narratives.med.wc_colors <- color_set_16()

narratives.med.wc <- narratives %>% 
  dplyr::filter(!is.na(response)) %>% 
  mutate(.responseWordCount = word_count(response)) %>%
  mutate(Month = as.Date(paste0(Month, "-01"), format = "%Y-%m-%d"))
```

```{r partner-over-time, include=FALSE}
narratives.med.wc_byPartner <- narratives.med.wc %>% 
  group_by(partnerName, Month) %>%
  summarize(median = median(.responseWordCount)) %>% 
  ungroup()

narratives.med.wc_byPartner_ts <-
  narratives.med.wc_byPartner %>%
  mutate(Month = yearmonth(Month)) %>% 
  as_tsibble(index = "Month", key = "partnerName") %>% 
  fill_gaps(median = 0, .full = TRUE)
```

```{r, fig.width=10, fig.height=10, fig.align='center'}
narratives.med.wc_byPartner_ts %>%
  ggplot(aes(Month, median, color = partnerName)) +
  geom_line() +
  geom_point() +
  geom_point(shape = 21, color = "black") +
  labs(
    title = "Median word count of responses per partner over months",
    subtitle = ""
  ) +
  xlab("Month") +
  ylab("Median word count") +
  scale_x_date(labels = date_format("%Y-%m")) +
  scale_colour_manual(values = narratives.med.wc_colors) +
  facet_wrap( ~ partnerName, scales = "free_x", ncol = 2) +
  theme_ecuador1(border = TRUE) +
  guides(colour = FALSE) +
  ## y lim is a bit more than max median value:
  ylim(0, max(narratives.med.wc_byPartner_ts$median) + 50L) +
  xlab(NULL)
```

```{r partner-subPartner-over-time, include=FALSE}
narratives.med.wc_byPartnerSubPartner <- narratives.med.wc %>% 
  group_by(partnerName, subPartnerName, Month) %>%
  summarize(median = median(.responseWordCount)) %>% 
  ungroup()

narratives.med.wc_byPartnerSubPartner_ts <-
  narratives.med.wc_byPartnerSubPartner %>%
  mutate(Month = yearmonth(Month)) %>% 
  as_tsibble(index = "Month", key = c("partnerName", "subPartnerName")) %>% 
  fill_gaps(median = 0, .full = TRUE)
```

<br>

#### Median word count of responses per specific Reporting Partners over months {-}

```{r, specific-reporting-partners-over-months, fig.height=14, fig.width=8, fig.align='center'}
.med.wc.specific <- narratives.med.wc_byPartnerSubPartner_ts %>% 
  dplyr::filter(partnerName %in% c("ACNUR", "UNICEF") | subPartnerName %in% "HIAS")

.med.wc.specific.splitted <- split(
  x = .med.wc.specific,
  f = .med.wc.specific$partnerName
)

narratives.med.wc.specific.partners <- 
  lapply(seq_along(.med.wc.specific.splitted), function(i) {
      ggplot(.med.wc.specific.splitted[[i]], aes(Month, median, color = subPartnerName)) +
      facet_wrap( ~ partnerName, scales = "free") +
      geom_line() +
      geom_point() +
      geom_point(shape = 21, color = "black") +
      xlab(NULL) +
      ylab("Median word count") +
      scale_x_date(labels = scales::date_format("%Y-%m"), date_breaks = "1 month") +
      theme_ecuador1(border = TRUE) +
      theme(legend.title = element_blank()) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1.1))
  })
cowplot::plot_grid(plotlist = narratives.med.wc.specific.partners, 
                   nrow = length(narratives.med.wc.specific.partners))
```

