
# Most used bigrams {-}

```{r, include = FALSE}
n.most.bigram <- 40L
```

```{r}
most.used.bigrams <- response.tokens.bigrams %>% 
  count(bigram) %>% 
  arrange(desc(n)) %>% 
  slice(seq(n.most.bigram))
```

```{r}
create_most_bigram_matrix <- function(group.column) {
  exist <- response.tokens.bigrams %>% 
    group_by({{group.column}}) %>% 
    dplyr::filter(bigram %in% most.used.bigrams$bigram) %>% 
    count(bigram) %>% 
    rename(group = {{group.column}}) %>% 
    select(group, bigram, n) %>% 
    ungroup()
  
  exist.groups <- unique(exist$group)
  
  dummy <- do.call(rbind, lapply(exist.groups, function(g) {
    bgrams <- exist[exist$group == g, "bigram", drop = TRUE]
    diff.bigrams <- setdiff(most.used.bigrams$bigram, bgrams)
    tibble::tibble(group = g, bigram = diff.bigrams, n = 0)
  }))
  
  ret <- rbind(exist, dummy)
  ret
}
```

```{r, include=FALSE}
bigram.chosen.formNames <- c("Tecnico", "Poblacion", "VBG")
```

Top *`r n.most.bigram`* bigrams are selected by cumulative use. The sectors *`r mmy::list_en_end(bigram.chosen.formNames)`* are chosen for the
visualization below of the most used bigrams because these sectors are the ones
with higher word counts (see [**this section**](#average-word-count-per-sector) for the
reason).

<br>

```{r, fig.height=8, fig.width=8, fig.align="center", fig.cap="Most common bigrams in different sectors"}
create_most_bigram_matrix(group.column = formNameRecode) %>% 
  dplyr::filter(group %in% bigram.chosen.formNames) %>% 
  ggplot(aes(group, bigram)) + 
  geom_tile(aes(fill = n), colour = "white") + 
  scale_fill_gradient(low = "white", high = "steelblue") +
  theme_light() +
  labs(
    title = "**What are the most used bigrams from responses in sectors?**",
    subtitle = paste0("The sectors in *", paste(bigram.chosen.formNames, collapse = ", "), "*"),
    caption = RECODE.TABLE.CAPTION
  ) +
  xlab(NULL) +
  ylab(NULL) +
  theme(
    plot.title = ggtext::element_markdown(lineheight = 1.1),
    plot.subtitle = ggtext::element_markdown(lineheight = 1.1)
  )
```

<br>

Below some more extra visualizations with location data. Unfortunately, the
bigram matrix plot did not bring interesting results when used with the partners
data so they are not included here.

<br>

```{r, fig.align="center"}
create_most_bigram_matrix(group.column = province) %>% 
  ggplot(aes(group, bigram)) + 
  geom_tile(aes(fill = n), colour = "white") + 
  scale_fill_gradient(low = "white", high = "orange") +
  theme_light() +
  theme(axis.text.x = element_text(
    angle = 330,
    vjust = 0.8,
    hjust = 0.2,
    color = "grey50"
  )) +
  labs(
    title = "What are the most used bigrams from responses in provinces?"
  ) +
  xlab(NULL) +
  ylab(NULL)
```

<br>

```{r, fig.align="center"}
create_most_bigram_matrix(group.column = canton) %>% 
  ggplot(aes(group, bigram)) + 
  geom_tile(aes(fill = n), colour = "white") + 
  scale_fill_gradient(low = "white", high = "red") +
  theme_light() +
  theme(axis.text.x = element_text(angle=330, vjust = 0.8, hjust = 0.2, color = "grey50")) +
  labs(
    title = "What are the most used bigrams from responses in cantons?"
  ) +
  xlab(NULL) +
  ylab(NULL)
```

<br>

Lastly, the table below shows the most used bigrams in all narrative responses
that is without any aggregation by any group such as sector name, location and
so on.

```{r}
most.used.bigrams %>% 
  gt() %>% 
  gt_condensed_style() %>% 
  cols_align(
    align = "left",
    columns = vars(bigram)
  ) %>% 
  cols_align(
    align = "center",
    columns = vars(n)
  ) %>% 
  tab_header(
    title = md("**Most used bigrams in all responses**"),
    subtitle = "The number of bigrams in the multi-text (narrative) fields"
  )
```

