
# The number of narrative records in sectors {-}

Since a robust design leads to better planning and monitoring
system, it is vital to detect and understand which information is not filled by
users.

If some sectors do not receive enough multi-line text data, it could be a very
well indicator to make educated guesses about some things do not go well in the
sectors when comes to enter long texts.

Surely, it is always better to find out the reason 'why'; however, in some cases,
removing the narrative fields from the sectors, which do not receive enough
multi-text data, might also be a meaningful move.

```{r}
narrative.all.records.tbl <- inner_join({
  narratives %>% 
    group_by(formNameRecode) %>% 
    summarise(filled = sum(!is.na(response)),
              missing = sum(is.na(response)))
},{
  narratives %>% 
    group_by(formNameRecode) %>% 
    summarise(total = length(response))
}, by = "formNameRecode") %>%
  group_by(formNameRecode) %>% 
  mutate(`Response rate` = round(filled/missing, 3)) %>% 
  arrange(desc(filled)) %>% 
  ungroup()

n.a.r.t <- inner_join(recode_tbl, narrative.all.records.tbl, by = "formNameRecode")

narrative.all.records.tbl %>% 
  gt() %>% 
  gt_condensed_style() %>%
  cols_label(
    formNameRecode = "Sector name"
  ) %>% 
  tab_spanner(
    label = "Response",
    columns = vars(filled, missing, total)
  ) %>% 
  tab_footnote(
    footnote = md("*Divide the number of `filled` responses in the number of `missing` responses*"),
    locations = cells_column_labels(
      columns = vars(`Response rate`)
    )
  )
```

<br>

Note that we also keep the narratives that are empty (which are displayed as
`NA`, *Not Available*).

The table shows the narrative fields with count of *'full*' and '*empty*' fields.

```{r, results='asis'}
ss.n.a.r.t <- n.a.r.t %>% arrange(desc(total)) %>% slice(seq(5))
for (i in seq(nrow(ss.n.a.r.t))) {
  ss <- ss.n.a.r.t[i, ]
  txt <- glue::glue(
    "**{formNameRecode}** (*{formName}*) has *{filled}* and *{missing}* records in a total of *{total}* records with a response rate of *{responseRate}*.",
    formNameRecode = ss[["formNameRecode"]],
    formName = ss[["formName"]],
    filled = ss[["filled"]],
    missing = ss[["missing"]],
    total = ss[["total"]],
    responseRate = ss[["Response rate"]]
  )
  cat("+", txt, "\n\n")
}
```

```{r, include=FALSE}
response.rate.threshold <- 0.1

```

On the other hand, the following sectors have response rate less than *`r
response.rate.threshold`* so they have very low response rate compared to all
sectors in the database.

```{r}
small.n.a.r.t. <- n.a.r.t[n.a.r.t$`Response rate` < response.rate.threshold, ]
for (i in seq(nrow(small.n.a.r.t.))) {
  ss <- small.n.a.r.t.[i, ]
  txt <- glue::glue(
    "**{formNameRecode}** (*{formName}*) has *{filled}* and *{missing}* records in a total of *{total}* records with a response rate of *{responseRate}*.",
    formNameRecode = ss[["formNameRecode"]],
    formName = ss[["formName"]],
    filled = ss[["filled"]],
    missing = ss[["missing"]],
    total = ss[["total"]],
    responseRate = ss[["Response rate"]]
  )
  cat("+", txt, "\n\n")
}
```

Please note that the following sectors are not included in the table above
because they do not contain any narrative fields in their designed forms:
`r mmy::list_en_end(paste0("*",setdiff(unique(form.table$formName), unique(narratives$formName)),"*"),add.quotes=T)`

