
# The number of narrative records in sectors {-}

```{r}
narrative.all.records.tbl <- inner_join({
  narratives %>% 
    group_by(formNameRecode) %>% 
    summarise(filled = sum(!is.na(response)),
              missing = sum(is.na(response)))
},{
  narratives %>% 
    group_by(formNameRecode) %>% 
    summarise(total = length(response))
}, by = "formNameRecode") %>%
  group_by(formNameRecode) %>% 
  mutate(`Response rate` = round(filled/missing, 3)) %>% 
  arrange(desc(filled)) %>% 
  ungroup()

n.a.r.t <- inner_join(recode_tbl, narrative.all.records.tbl, by = "formNameRecode")

narrative.all.records.tbl %>% 
  gt() %>% 
  gt_condensed_style() %>%
  cols_label(
    formNameRecode = "Sector name"
  ) %>% 
  tab_spanner(
    label = "Response",
    columns = vars(filled, missing, total)
  ) %>% 
  tab_footnote(
    footnote = md("*Divide the number of `filled` responses in the number of `missing` responses*"),
    locations = cells_column_labels(
      columns = vars(`Response rate`)
    )
  )
```

Note that we also keep the narratives that are empty (which are displayed as
`NA`, *Not Available*).

<br>

The table shows the narrative fields with count of *'full*' and '*empty*' fields.

```{r, results='asis'}
ss.n.a.r.t <- n.a.r.t %>% arrange(desc(total)) %>% slice(seq(5))
for (i in seq(nrow(ss.n.a.r.t))) {
  ss <- ss.n.a.r.t[i, ]
  txt <- glue::glue(
    "The form **{formNameRecode}** (*{formName}*) has *{filled}* and *{missing}* records in a total of *{total}* records.",
    formNameRecode = ss[["formNameRecode"]],
    formName = ss[["formName"]],
    filled = ss[["filled"]],
    missing = ss[["missing"]],
    total = ss[["total"]]
  )
  cat("+", txt, "\n\n")
}
```

Please note that the following sectors are not included in the table above
because they do not contain any narrative fields:
`r mmy::list_en_end(paste0("*",setdiff(unique(form.table$formName), unique(narratives$formName)),"*"),add.quotes=T)`

