# Partners {-#partners}

They are two types of partners in the database:

+ **Reporting partners**: Higher level of partners reporting directly in *ActivityInfo*.

+ **Implementing partners**: Partners reporting through a reporting partner.

<br>

```{r, include=FALSE}
ppartners <- local({
  p <- form.table %>%
    group_by(recordId, partnerName) %>% 
    count(partnerName)
  pc <- p %>% 
    group_by(partnerName) %>% 
    count(partnerName, name = "freq", sort = TRUE)
  perc <- pc %>% 
    ungroup() %>%
    mutate(prop = round(freq / sum(freq), 3))
})
```

```{r}
ppartners %>%
  rename(
    `Reporting partner` = partnerName,
    `Frequency` = freq,
    `Relative frequency` = prop
  ) %>% 
  gt() %>% 
  cols_align(
    align = "left",
    columns = vars(`Reporting partner`)
  ) %>% 
  cols_align(
    align = "center",
    columns = vars(`Frequency`, `Relative frequency`)
  ) %>% 
  tab_header(
    title = md("**Reporting partners in all field types**")
  ) %>%
  gt_condensed_style()
```

The table above shows the count of reporting partner per each record:

+ **`r ppartners[1,"partnerName",drop=T]`** has *`r xfun::n2w(ppartners[1,"freq",drop=T])`*
records, 
which is *`r scales::percent(ppartners[1,"prop",drop=T])`*
of the total records.

+ Second, 
**`r ppartners[2,"partnerName",drop=T]`** has *`r xfun::n2w(ppartners[2,"freq",drop=T])`*
records, 
which is *`r scales::percent(ppartners[2,"prop",drop=T])`*
of the total records.

<br>

```{r, include=FALSE}
ppartnersAndSub <- local({
  pSub <- form.table %>%
    group_by(recordId, partnerName, subPartnerName) %>% 
    count(partnerName)
  pcSub <- pSub %>% 
    group_by(partnerName, subPartnerName) %>% 
    count(partnerName, subPartnerName, name = "freq", sort = TRUE)
  percSub <- pcSub %>% 
    summarise(freq = sum(freq)) %>% 
    mutate(prop = round(freq / sum(freq), 3))
})
```

```{r, include=FALSE}
ppartnersAndSub.acnur.hias <- ppartnersAndSub %>% 
  dplyr::filter(partnerName %in% "ACNUR", subPartnerName %in% "HIAS")
```

```{r, echo=FALSE, warning=FALSE}
## Used parse & eval because of this issue: 
## https://github.com/rstudio/gt/issues/308
ppartnersAndSub.bold_groups <- glue::glue_collapse(glue::glue('
  tab_style(
    locations = cells_group(groups = "{x}"),
    style = cell_text(weight = "bold")
  )
  ',
  x = unique(ppartnersAndSub$partnerName)
  ),
  sep = " %>%\n")
ppartnersAndSub.tbl <- ppartnersAndSub %>% 
  group_by(subPartnerName) %>% 
  arrange(desc(freq)) %>% 
  group_by(partnerName) %>% 
  rename(
    `Reporting partner` = partnerName,
    `Implementing partner` = subPartnerName,
    `Frequency` = freq,
    `Relative frequency` = prop
  )
eval(parse(text=glue::glue('
ppartnersAndSub.tbl %>%
  gt() %>% 
  cols_align(
    align = "left",
    columns = vars(`Implementing partner`)
  ) %>% 
  cols_align(
    align = "center",
    columns = vars(`Frequency`, `Relative frequency`)
  ) %>% 
  fmt_missing(
    columns = vars(`Implementing partner`),
    missing_text = md("*NA*")
  ) %>%
  tab_header(
    title = md("**Reporting & implementing partners in all field types**")
  ) %>%
  gt_condensed_style() %>% 
  {bolds}
', bolds = ppartnersAndSub.bold_groups)))
```

The table above shows the proportion of records entered by
reporting and implementing partners.

+ *`r ppartnersAndSub.acnur.hias[,"freq",drop=T]`* out of 
*`r sum(ppartnersAndSub %>% dplyr::filter(partnerName %in% "ACNUR") %>% pull(freq))`* 
total responses 
(*`r scales::percent(ppartnersAndSub.acnur.hias[["prop"]])`*)
of **ACNUR** is actually coming from **HIAS**.

+ **UNICEF** has more diversed partners in terms of reporting.
*`r scales::percent(ppartnersAndSub%>%dplyr::filter(partnerName%in% "UNICEF", subPartnerName %in% "HIAS")%>%pull(prop))`*
of responses of **UNICEF** comes from **HIAS**.
*`r scales::percent(ppartnersAndSub%>%dplyr::filter(partnerName %in% "UNICEF", subPartnerName %in% "UNICEF")%>% pull(prop))`*
of reporting comes from the **UNICEF** itself.

+ Under **PMA**, there are *`r nrow(ppartnersAndSub%>%dplyr::filter(partnerName %in% "PMA"))`* sub-partners. **HIAS** reports *`r scales::percent(ppartnersAndSub%>%dplyr::filter(partnerName %in% "PMA", subPartnerName %in% "HIAS")%>%pull(prop))`* of these records.

#### Partners filling Narrative (multi-line text) fields in records {-}

```{r, include=FALSE}
ppartners_narratives <- local({
  p <- narratives %>%
    group_by(recordId, partnerName) %>% 
    count(partnerName)
  pc <- p %>% 
    group_by(partnerName) %>% 
    count(partnerName, name = "freq", sort = TRUE)
  perc <- pc %>% 
    ungroup() %>%
    mutate(prop = round(freq / sum(freq), 3))
})
```

```{r}
ppartners_narratives %>%
  rename(
    `Reporting partner` = partnerName,
    `Frequency` = freq,
    `Relative frequency` = prop
  ) %>% 
  gt() %>% 
  cols_align(
    align = "left",
    columns = vars(`Reporting partner`)
  ) %>% 
  cols_align(
    align = "center",
    columns = vars(`Frequency`, `Relative frequency`)
  ) %>% 
  tab_header(
    title = md("**Reporting partners in *Narrative* field type**")
  ) %>%
  gt_condensed_style()
```

The table above shows the count of reporting partner in *Narrative* field per each record:

+ **`r ppartners_narratives[1,"partnerName",drop=T]`** has *`r xfun::n2w(ppartners_narratives[1,"freq",drop=T])`*
records, 
which is *`r scales::percent(ppartners_narratives[1,"prop",drop=T])`*
of the total records.

+ Second, 
**`r ppartners_narratives[2,"partnerName",drop=T]`** has *`r xfun::n2w(ppartners_narratives[2,"freq",drop=T])`*
records, 
which is *`r scales::percent(ppartners_narratives[2,"prop",drop=T])`*
of the total records.

<br>

```{r, include=FALSE}
ppartnersAndSub_narratives <- local({
  pSub <- narratives %>%
    group_by(recordId, partnerName, subPartnerName) %>% 
    count(partnerName)
  pcSub <- pSub %>% 
    group_by(partnerName, subPartnerName) %>% 
    count(partnerName, subPartnerName, name = "freq", sort = TRUE)
  percSub <- pcSub %>% 
    summarise(freq = sum(freq)) %>% 
    mutate(prop = round(freq / sum(freq), 3))
})
```

```{r, include=FALSE}
ppartnersAndSub_narratives.acnur.hias <- ppartnersAndSub_narratives %>% 
  dplyr::filter(partnerName %in% "ACNUR", subPartnerName %in% "HIAS")
```

```{r, echo=FALSE, warning=FALSE}
## Used parse & eval because of this issue: 
## https://github.com/rstudio/gt/issues/308
ppartnersAndSub_narratives.bold_groups <- glue::glue_collapse(glue::glue('
  tab_style(
    locations = cells_group(groups = "{x}"),
    style = cell_text(weight = "bold")
  )
  ',
  x = unique(ppartnersAndSub_narratives$partnerName)
  ),
  sep = " %>%\n")
ppartnersAndSub_narratives.tbl <- ppartnersAndSub_narratives %>% 
  group_by(subPartnerName) %>% 
  arrange(desc(freq)) %>% 
  group_by(partnerName) %>% 
  rename(
    `Reporting partner` = partnerName,
    `Implementing partner` = subPartnerName,
    `Frequency` = freq,
    `Relative frequency` = prop
  )
eval(parse(text=glue::glue('
ppartnersAndSub_narratives.tbl %>%
  gt() %>% 
  cols_align(
    align = "left",
    columns = vars(`Implementing partner`)
  ) %>% 
  cols_align(
    align = "center",
    columns = vars(`Frequency`, `Relative frequency`)
  ) %>% 
  fmt_missing(
    columns = vars(`Implementing partner`),
    missing_text = md("*NA*")
  ) %>%
  tab_header(
    title = md("**Reporting & implementing partners in in *Narrative* field type**")
  ) %>%
  gt_condensed_style() %>% 
  {bolds}
', bolds = ppartnersAndSub_narratives.bold_groups)))
```

The table above shows the proportion of records in *Narrative* field entered by
reporting and implementing partners.

+ *`r ppartnersAndSub_narratives.acnur.hias[,"freq",drop=T]`* out of 
*`r sum(ppartnersAndSub_narratives %>% dplyr::filter(partnerName %in% "ACNUR") %>% pull(freq))`* 
total responses 
(*`r scales::percent(ppartnersAndSub_narratives.acnur.hias[["prop"]])`*)
of **ACNUR** is actually coming from **HIAS**.

+ **UNICEF** has more diversed partners in terms of reporting.
*`r scales::percent(ppartnersAndSub_narratives%>%dplyr::filter(partnerName%in% "UNICEF", subPartnerName %in% "HIAS")%>%pull(prop))`*
of responses of **UNICEF** comes from **HIAS**.
*`r scales::percent(ppartnersAndSub_narratives%>%dplyr::filter(partnerName %in% "UNICEF", subPartnerName %in% "UNICEF")%>% pull(prop))`*
of reporting comes from the **UNICEF** itself.

+ Under **PMA**, there are *`r nrow(ppartnersAndSub_narratives%>%dplyr::filter(partnerName %in% "PMA"))`* sub-partners. **HIAS** reports *`r scales::percent(ppartnersAndSub_narratives%>%dplyr::filter(partnerName %in% "PMA", subPartnerName %in% "HIAS")%>%pull(prop))`* of these records.

<br> 

```{r, include=FALSE}
partners_compare_text_func <- function(tbl1, tbl2, partner.name) {
  narr <- tbl1[tbl1$partnerName==partner.name,"freq",drop=T]
  all <- tbl2[tbl2$partnerName==partner.name,"freq",drop=T]
  paste0(round(100*narr/all, 2), "%")
}
```

+ Additionally, *`r partners_compare_text_func(ppartners_narratives, ppartners, "ACNUR")`* 
  of the records have *Narrative* fields in **ACNUR**.

