# Partners entering narrative data {-}

Here, we look at the partners entering narrative data. The number of missing
records (technical saying: `NA`s) are excluded.

```{r}
narratives.part.count <- narratives %>% 
  dplyr::filter(!is.na(response)) %>% ## remove NAs
  group_by(partnerName, subPartnerName) %>% 
  count(partnerName, subPartnerName, name = "freq", sort=TRUE) %>% 
  summarize(freq = sum(freq)) %>% 
  mutate(prop = round(freq/sum(freq), 3)) %>% 
  mutate(percent = paste0(floor(100 * freq/sum(freq)), "%"))

## partners having multiple sub-partners
narratives.part.count.with.subPartners <- narratives.part.count %>% 
  dplyr::filter(prop != 1) %>% 
  ungroup()

## partners do not have any sub-partners reporting via them.
narratives.part.count.without.subPartners <- narratives.part.count %>% 
  dplyr::filter(prop == 1) %>% 
  ungroup()
```

As we have seen previously, Not all Reporting and Implementing Partners record
multi-line narrative textual data. For instance, the partner **PMA** has lots of
Implementing Partners reporting for the different data types (as seen in above)
but there are no narratives from them.

```{r, warning=FALSE, fig.width=7, fig.height=8, fig.align='center', fig.cap='"Reporting" and "Implementing" partners reporting narrative data'}
## grouped facet wraps. Source:
## https://stackoverflow.com/questions/14840542/place-a-legend-for-each-facet-wrap-grid-in-ggplot2

plot_partners_splitted <- split(
  x = narratives.part.count.with.subPartners, 
  f = narratives.part.count.with.subPartners$partnerName
)

## bring ACNUR and UNICEF plots to the front:
plot_partners_splitted_front <- plot_partners_splitted[c(
  c("ACNUR", "UNICEF"),
  setdiff(names(plot_partners_splitted), c("ACNUR", "UNICEF"))
)]

plot_partners_list <- lapply(seq_along(plot_partners_splitted_front), function(i) {
  ggplot(plot_partners_splitted_front[[i]], aes(x = partnerName, y = prop)) +
    geom_bar(aes(fill = reorder(subPartnerName,-prop)), position = "dodge", stat =
               "identity", color = "black") +
    facet_wrap(~ partnerName, scales = "free") +
    geom_label(
      aes(
        label = percent,
        group = reorder(subPartnerName,-prop),
        vjust = 0.65
      ),
      position = position_dodge(width = 1),
      label.size = 0.35
    ) +
    ylim(0, 1) + 
    xlab(NULL) +
    ylab(NULL) +
    theme_ecuador1(border = TRUE) +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      legend.title = element_blank(),
      legend.position = "bottom"
    ) +
    scale_fill_brewer(palette = eval(quote(brewer_diverging_color_names()[i])))
})

## "cowplot" is civilized version of 'gridExtra'.
do.call(cowplot::plot_grid, plot_partners_list)
```

+ *`r subset(narratives.part.count.with.subPartners, partnerName == "ACNUR" & subPartnerName == "HIAS", "percent", drop = T)`*
of the narrative records are
entered by the implementing partner **HIAS** reported via **ACNUR**.
Only
*`r subset(narratives.part.count.with.subPartners, partnerName == "ACNUR" & subPartnerName == "ACNUR", "percent", drop = T)`*
of the narrative records are entered by **ACNUR** itself.

+ Also **HIAS** enters *`r subset(narratives.part.count.with.subPartners, partnerName == "UNICEF" & subPartnerName == "HIAS", "percent", drop = T)`* of the narrative records via **UNICEF**.

+ The rest of the "reporting partners" do not have any "implementing partners",
as it seems that they do the implementation:
`r mmy::list_en_end(paste0("*",unique(narratives.part.count.without.subPartners$partnerName),"*"),add.quotes=T)`.

