
# "Qualitativo" questions {-}

*Do the questions formed by the word "qualitativo" create lesser or longer responses?*

Almost all questions have a form of the term "qualitative" in the question
fields. Here we are interested in the questions that they have nothing but the
term "qualitative".

```{r, include=FALSE}
cual_matches <- local({
  uniq <- unique(narratives.wc$question)
  cual <- "^[C|c]ualitativo$"
  qual <- "^[Q|q]ualitative$"
  act <- "^[A|a]ct_[[:alnum:]]_[C|c]ualitativo$"
  unique(c(
    uniq[grep(cual, uniq)],
    uniq[grep(qual, uniq)],
    uniq[grep(act, uniq)]
  ))
})
```

```{r}
tibble::as_tibble(as.data.frame(t(cual_matches))) %>% 
  arrange(.[[1]]) %>% 
  gt() %>% 
  gt_condensed_style() %>% 
  cols_align(
    align = "center"
  )
```

<br/>

```{r}
## narratives with 'qualitative' question:
qual_narr_wc <- narratives.wc %>% 
  dplyr::filter(question %in% cual_matches)

## narratives without 'qualitative' question:
non_qual_narr_wc <- narratives.wc %>% 
  anti_join(qual_narr_wc, by = c("folderName", "formNameRecode", "Month", "recordId"))
```

#### Compare the responses with which has a question called qualitative (and variants) with the rest of the responses {-}

We look at the specific sectors such as **VBG**, **Poblacion** and **Tecnico**
as they have higher sample sizes. This situation is explained in this
[**section**](#average-word-count-per-sector).

```{r}
plot_narr_qual <- function(data, 
                           plot.title, plot.subtitle,
                           plot.fill.color.palette) {
  {{data}} %>% 
    filter(formNameRecode %in% c("VBG", "Poblacion", "Tecnico")) %>%
    ggplot(aes(formNameRecode, .responseWordCount, fill = formNameRecode)) +
    geom_boxplot(outlier.fill = "orange", outlier.color = "red", outlier.shape = 23, outlier.alpha = 1) +
    geom_jitter(alpha = 0.4) +
    coord_flip() +
    labs(
      title = plot.title,
      subtitle = plot.subtitle,
      caption = RECODE.TABLE.CAPTION
    ) +
    xlab(NULL) +
    ylab(NULL) +
    guides(fill = FALSE) +
    theme_ecuador1() +
    theme(
      plot.title = ggtext::element_markdown(lineheight = 1.1),
      plot.subtitle = ggtext::element_markdown(lineheight = 1.1)
    ) +
    ggplot2::scale_fill_brewer(type = "div", palette = plot.fill.color.palette)
}
```

```{r}
plot_narr_qual(
  qual_narr_wc,
  plot.title = "Response word counts by sector ***with 'qualitative'***",
  plot.subtitle = paste(
    "A box plot distribution where individual points show only the responses for which the responses",
    "have a 'qualitative' (or a variant) word in its question",
    sep = "<br>"
  ),
  plot.fill.color.palette = 1
)
```

```{r}
plot_narr_qual(
  non_qual_narr_wc,
  plot.title = "Response word counts by sector ***without 'qualitative'***",
  plot.subtitle = paste(
    "A box plot distribution where individual points show only the responses for which the responses",
    "don't have any 'qualitative' (or a variant) word in its question",
    sep = "<br>"
  ),
  plot.fill.color.palette = 5
)
```

We can perform an independent *t-test* on these sample data to determine whether
there is a statistically significant difference between the means in word count
of these groups.

The sample data we have at hand is that:

1. Responses with which its question has a 'qualitative' or a variant word in it,

2. Responses with whose question doesn't have any 'qualitative' or variant
words.

Therefore, we can define the null and alternative hypotheses as that:

$$H_0 : \mu_1 = \mu_2$$
$$H_A : \mu_1 \neq \mu_2$$

The *t-test* has an assumption that the *sample* data must be normally distributed.

First, we look at the distribution of the sample data, for both responses.

```{r}
## rbind unequal data.frames with `merge + all = TRUE` combination
qual_resp_rbind <- as_tibble(merge(
  tibble(qualitative_resp_wc = qual_narr_wc$.responseWordCount),
  tibble(non_qualitative_resp_wc = non_qual_narr_wc$.responseWordCount),
  by = "row.names",
  all = TRUE
)) %>%
  select(-Row.names)

qual_resp_rbind %>% 
  inspect_num() %>% 
  show_plot()
```

The distribution in the first group looks less variant than the second one.

Second, we observe the QQ test:

```{r}
qqnorm(qual_resp_rbind$qualitative_resp_wc)
qqnorm(qual_resp_rbind$non_qualitative_resp_wc)
```

We can additionally run the Shapiro–Wilk normality test.

```{r}
bind_rows(
  bind_cols(
    tibble(sample = "Qualitative"), broom::tidy(shapiro.test(qual_resp_rbind$qualitative_resp_wc))
  ),
  bind_cols(
    tibble(sample = "Non-qualitative"), broom::tidy(shapiro.test(qual_resp_rbind$non_qualitative_resp_wc))
  )
) %>% 
  gt()
```

```{r}
t.test(qual_resp_rbind$qualitative_resp_wc, qual_resp_rbind$non_qualitative_resp_wc)
```

It seems that we fail to reject null the hypothesis *at the p < 0.05 level* that
there is no statistically significant difference between means of these two
samples.

