
# The "Qualitativo" questions {-}

#### Do the responses have a question consisted of the word "qualitativo" lead to lesser word counts than the others? {-}

Every response field in *ActivityInfo* forms has a question field. Particularly,
in the Ecuador monitoring database, some responses have question field, which is
only filled as one-word length "qualitative" or similar variants, and the other
responses have different question fields, which seem to look like that they are
formed as proper questions.

Let's say that we are interested in that there are any meaningful differences
between the word counts of these responses: one the one hand, the responses
whose question field has "qualitative", and on the other hand, the remaining
responses.

```{r, include=FALSE}
cual_matches <- local({
  uniq <- unique(narratives.wc$question)
  cual <- "^[C|c]ualitativo$"
  qual <- "^[Q|q]ualitative$"
  act <- "^[A|a]ct_[[:alnum:]]_[C|c]ualitativo$"
  unique(c(
    uniq[grep(cual, uniq)],
    uniq[grep(qual, uniq)],
    uniq[grep(act, uniq)]
  ))
})
```

The variants of "qualitative" term in the question fields are as follows:
`r mmy::list_en_end(paste0("*", sort(cual_matches), "*"), add.quotes = TRUE)`

```{r}
## narratives with 'qualitative' question:
qual_narr_wc <- narratives.wc %>% 
  dplyr::filter(question %in% cual_matches)

## narratives without 'qualitative' question:
non_qual_narr_wc <- narratives.wc %>% 
  anti_join(qual_narr_wc, by = c("folderName", "formNameRecode", "Month", "recordId"))
```

For this kind of analysis, we look at specific sectors such as **VBG**,
**Poblacion** and **Tecnico** because of a couple of reasons. First, they have
higher and similar sample sizes compared to the other sectors in the database.
Second, their variances are similar. The rationale behind this choice is
explained in the [**section**](#average-word-count-per-sector).

```{r}
plot_narr_qual <- function(data, 
                           plot.title, plot.subtitle,
                           plot.fill.color.palette) {
  {{data}} %>% 
    filter(formNameRecode %in% c("VBG", "Poblacion", "Tecnico")) %>%
    ggplot(aes(formNameRecode, .responseWordCount, fill = formNameRecode)) +
    geom_boxplot(outlier.fill = "orange", outlier.color = "red", outlier.shape = 23, outlier.alpha = 1) +
    geom_jitter(alpha = 0.4) +
    coord_flip() +
    labs(
      title = plot.title,
      subtitle = plot.subtitle,
      caption = RECODE.TABLE.CAPTION
    ) +
    xlab(NULL) +
    ylab(NULL) +
    guides(fill = FALSE) +
    theme_ecuador1() +
    theme(
      plot.title = ggtext::element_markdown(lineheight = 1.1),
      plot.subtitle = ggtext::element_markdown(lineheight = 1.1)
    ) +
    ggplot2::scale_fill_brewer(type = "div", palette = plot.fill.color.palette)
}
```

```{r}
plot_narr_qual(
  qual_narr_wc,
  plot.title = "Response word counts by sector ***with 'qualitative'***",
  plot.subtitle = paste(
    "A box plot distribution where individual points show only the responses for which the responses",
    "have a 'qualitative' (or a variant) word in its question",
    sep = "<br>"
  ),
  plot.fill.color.palette = 1
)
```

```{r}
plot_narr_qual(
  non_qual_narr_wc,
  plot.title = "Response word counts by sector ***without 'qualitative'***",
  plot.subtitle = paste(
    "A box plot distribution where individual points show only the responses for which the responses",
    "don't have any 'qualitative' (or a variant) word in its question",
    sep = "<br>"
  ),
  plot.fill.color.palette = 5
)
```

#### Comparison of two groups by their means {-}

After all, we have two groups (which they can also be called as 
*random processes* due to their distinct characteristics) that are from these
sectors:

1. **Qualitative group**: Responses with which their questions have a
'qualitative' or a variant word.

2. **Non-qualitative group**: Responses with which their questions do not have
any 'qualitative' or any variant words.

To begin with, we try to identify the underlying distribution of these groups:

```{r}
## rbind unequal data.frames with `merge + all = TRUE` combination
qual_resp_cbind <- as_tibble(merge(
  tibble(qualitative_resp_wc = qual_narr_wc$.responseWordCount),
  tibble(non_qualitative_resp_wc = non_qual_narr_wc$.responseWordCount),
  by = "row.names",
  all = TRUE
)) %>%
  select(-Row.names)

qual_resp_cbind %>% 
  inspect_num() %>% 
  show_plot()
```

In the histograms above, we can clearly see that the distribution in the first
group looks less deviated than the distribution of the second one.

In addition to the histograms, we can generate the *Q-Q (quantile-quantile)
plots* to assess from a different way whether the grouped data deviates from
normal:

```{r}
qq_plot <- function(x, title) {
  qqnorm(x, main = title)
  qqline(x)
}
par(mfrow=c(1,2))
qq_plot(qual_resp_cbind$qualitative_resp_wc, "Qualitative group")
qq_plot(qual_resp_cbind$non_qualitative_resp_wc, "Non-qualitative group")
```

The first Q-Q plot shows that the deviations from the straight line are high.
This should indicate that the data does not deviate from normal. The
observations in the second plot are more aligned with its line.

```{r,include=FALSE}
N_bootstrap <- 1000
```

Since the assumption of normality is violated in these *random processes*, we
consider performing *bootstrapping*, which resamples the data with replacements
so that the observations can be selected more than once, 
with *`r N_bootstrap`* samples. After that, we can use the bootstrapped data to
calculate the confidence intervals.

```{r, include=FALSE}
library(boot)
mean_difference <- function(df, cols, indices) {
  m1 <- mean(df[indices, cols[1], drop=TRUE], na.rm = TRUE)
  m2 <- mean(df[indices, cols[2], drop=TRUE], na.rm = TRUE)
  m1 - m2
}
results <- boot(
  as.data.frame(qual_resp_cbind),
  statistic = mean_difference,
  cols = c("qualitative_resp_wc", "non_qualitative_resp_wc"),
  R = N_bootstrap
)
```

After bootstrapping, the distribution of groups looks pretty normal:

```{r}
plot(results)
```

We calculate the *bootstrap confidence intervals* at the confidence levels of
*90*, *95* and *99* in order to see the difference between the means in response
word count of these groups.

```{r}
(ci <- boot.ci(results, conf = c(0.90, 0.95, 0.99), type="bca"))
```

As a result, based on the resampled data, the difference in mean between those
groups at *95%* confidence level are 
in between *`r round(ci$bca[2,4], 2)`* and *`r round(ci$bca[2,5], 2)`*. 
A crude interpretation of this would be that the mean interval does not show too
much significance that the terms in the question fields do not affect the word
count in the response fields. A further either quantiative or qualitative study
is recommended.

