
# Which sectors do receive high word counts in responses? {-}

A rule of thumb is to know the distributions separated by the major category,
which is here the sector name.

```{r plot-topic-response-wc, fig.width=8, fig.height=7, fig.align='center'}
narratives.wc %>%
  mutate(formNameRecode = fct_reorder(formNameRecode, .responseWordCount)) %>% 
  ggplot(aes(formNameRecode, .responseWordCount, fill = formNameRecode)) +
  geom_boxplot(outlier.fill = "orange", outlier.color = "red", outlier.shape = 23, outlier.alpha = 1) +
  coord_flip() +
  labs(
    title = "Response and word counts by sector",
    subtitle = "A box plot distribution showing the variance of word counts in different sectors",
    caption = RECODE.TABLE.CAPTION
  ) +
  scale_y_continuous(labels = function(x) paste(x, "words")) +
  xlab(NULL) +
  ylab(NULL) +
  guides(fill = FALSE) +
  theme_ecuador1() 
```

Box plots are commonly used to visualize the *measure of spread* in the data
that gives a clear picture about how the variability and dispersion of the raw
data look like.

The boxplot has some characteristics that should be known before diving into the
interpretation:

1) left and right borders of the central rectangle present the
first and third quartile values of interquartile range (IQR) respectively;

2) the line in the middle of the rectangle indicates the *median* value;

3) the end of the lines streched from both right and left sides of the central
rectangle point the maximum and minimum values;

4) the orange colored points show the *outliers* in the data.

Therefore, we see a number of things:

As we have seen in the previous sections talked about the variability and the
sample size of the data, the sectors are not equally distributed.

```{r, include=FALSE}
max.response.sector <- narratives.wc[which.max(narratives.wc$.responseWordCount), ]
max.sectors <- narrative_mean_diff[narrative_mean_diff$sampleSize > 100,]
```

+ The sector *`r max.response.sector$formNameRecode`* 
(or long name, *`r max.response.sector$formName`*)
has the longest response among all the responses that is 
*`r max.response.sector$.responseWordCount`* words 
(please see the recordId: *`r max.response.sector$recordId`* for more information).

+ We have looked the average word count in sample
sizes in each sector
([**see the section**](#average-word-count-per-sector)),
therefore the sectors
`r mmy::list_en_end(paste0("**", max.sectors$formNameRecode, "**"), add.quotes = TRUE)` 
have the highest sample size.

+ The graph shows that these sectors have several outliers, which are very long
texts compare to the average.

