
# Variance of response word count per sector {-}

If we look at the *standard deviation*, which is a statistics to show the
*measure of spread* in data, of word count of responses in each sector, we can clearly see that
some of the sectors have high *variability*.

That's important metric to consider which sectors do have responses that are
worth to conduct analyses.

<br>

```{r sd-form-topic-table}
rwc.sd <- narratives.wc %>% 
  group_by(formNameRecode) %>% 
  summarize(SD = sd(.responseWordCount))
rwc.sample <- narratives.wc %>% 
  group_by(formNameRecode) %>% 
  summarise(sample_size = length(.responseWordCount))
rwc.sd.sample.join <- rwc.sd %>%
  left_join(rwc.sample, by = "formNameRecode") %>% 
  arrange(desc(sample_size))
rwc.sd.sample.join %>% 
  gt() %>%
  fmt_number(
    columns = vars(SD),
    decimals = 2,
    use_seps = T
  ) %>%
  fmt_missing(
    columns = vars(SD),
    missing_text = md("*NA*")
  ) %>%
  cols_label(
    formNameRecode = "Sector name",
    sample_size = "Sample size",
  ) %>% 
  tab_header(
    title = "SD and Sample size of response word counts per sector name",
    subtitle = md("Measure the spread with *standard deviation*")
  )
```

The table above clearly shows that some of the sectors are quite different in
terms of the word count in responses. 

```{r}
top.high.sample.size <- rwc.sd.sample.join %>% arrange(desc(sample_size)) %>% slice(seq(3))
for (i in seq(nrow(top.high.sample.size))) {
  top <- top.high.sample.size[i, ]
  txt <- glue::glue(
    "**{formNameRecode}** has *{sample_size}* sample size with a standard deviation of *{SD}*",
    formNameRecode = top[["formNameRecode"]],
    sample_size = top[["sample_size"]],
    SD = signif(top[["SD"]], 4)
  )
   cat("+", txt, "\n\n")
}
```

On the other hand, some sectors have very low sample size and high deviation.

```{r sd-details-salud}
sd.details.salud <- narratives.wc %>% 
  dplyr::filter(formNameRecode == "Salud") %>% 
  select(.responseWordCount) %>% 
  pull(.responseWordCount)
```

+ **Salud** has
only *`r length(sd.details.salud)`* records with diverse word counts, 
which are *`r sd.details.salud`*.

