
# Diversion of response word count per sector {-}

If we look at the standard deviation, is a single number statistics to show the
*measure of spread* in data, of response word count per sector, we will see that
some of the sectors have high *variability*.

<br>

```{r sd-form-topic-table}
rwc.sd <- narratives.wc %>% 
  group_by(formNameRecode) %>% 
  summarize(SD = sd(.responseWordCount))
rwc.sample <- narratives.wc %>% 
  group_by(formNameRecode) %>% 
  summarise(sample_size = length(.responseWordCount))
rwc.sd %>%
  left_join(rwc.sample, by = "formNameRecode") %>% 
  arrange(desc(sample_size)) %>% 
  gt() %>%
  fmt_number(
    columns = vars(SD),
    decimals = 2,
    use_seps = T
  ) %>%
  fmt_missing(
    columns = vars(SD),
    missing_text = md("*NA*")
  ) %>%
  cols_label(
    formNameRecode = "Sector name",
    sample_size = "Sample size",
  ) %>% 
  tab_header(
    title = "SD and Sample size of response word counts per sector name",
    subtitle = md("Measure the spread with *standard deviation*")
  )
```

```{r sd-details-salud}
sd.details.salud <- narratives.wc %>% 
  dplyr::filter(formNameRecode == "Salud") %>% 
  select(.responseWordCount) %>% 
  pull(.responseWordCount)
```

The table above clearly shows that some of the sectors are quite different in
terms of the word count in responses. For instance, the sector **Salud** has
only *`r length(sd.details.salud)`* records with diverse word counts, 
which are *`r sd.details.salud`*.

