
# Textual data preparation {-}

Describe how to prepare textual data and what common steps are usually performed.

They are usually four steps involved in this process:

**1. Tokenization**

Tokenization means to split a text into tokens considered
meaningful units of text. A token can either be a word (and often it is) or a
group of words (such as *bigram*), or even a sentence that depends on the level
of analysis.

```{r}
response.tokens_orig <- narratives %>%
  select(-reportingUsers) %>% 
  dplyr::filter(!is.na(response)) %>% ## remove NAs
  unnest_tokens(word, response, token = "words", to_lower = TRUE)

question.tokens_orig <- narratives %>% 
  select(-reportingUsers) %>% 
  dplyr::filter(!is.na(response)) %>% ## remove NAs
  unnest_tokens(word, question, token = "words", to_lower = TRUE)
```

Perform stemming, which you bring words (nouns/verbs) back to base or infinitive
forms, will be the next step after tokenization, so we can get the essence of
words.

**2. Strip punctuation**

Punctuation is often not required in text analysis (unless a researcher wants to
tokenize the text based on a specific classifier *such as sentence tokens*);
therefore, they create noise.

**3. Convert text into lowercase**

When the text turned into lowercase, for instance, the words *respuesta* and
*Respuesta* will no longer be taken as different words.

**4. Exclude stopwords & numbers**

Stop words usually mean the most common words in a language that will bring no
significant results in analysis. They are overly distributed in the text and
they will not give so meaningful results itself. Stop-words are including
articles (*el/la*), conjunctions (*y*), pronouns (*yo/tú/etc.*) and so on.

In text mining, this process is usually done
after the text converted into lowercase so one does not have to provide stop
words including both lower and sentence case versions.

We have imported a list of Spanish stopwords data (source
[**here**](https://github.com/stopwords-iso/stopwords-es), and that's the
alternative for `stopwords_es` list from the
[corpus](https://cran.r-project.org/package=corpus) package) and perform a
*filtering join* returning tokens from textual data by excluding the words
listed in the stopwords. that only returns the tokens not listed in the
stopwords.

```{r}
es_stopwords <- get_es_stopwords()

response.tokens_stopw <- response.tokens_orig %>% 
  anti_join(es_stopwords, by = "word")

act_codes <- data.frame(word = unique(question.tokens_orig$word[grep("act\\_\\d+.*", question.tokens_orig$word)]), stringsAsFactors = FALSE)

question.tokens_stopw <- question.tokens_orig %>% 
  anti_join(es_stopwords, by = "word") %>% 
  anti_join(act_codes, by = "word")
```

The original tokens for the response originally have 
*`r nrow(response.tokens_orig)`*
rows. However, after merging stop words, the number of rows have decreased to
*`r nrow(response.tokens_stopw)`* and that the change in between is
*`r round(convolve(nrow(response.tokens_stopw), 100)/nrow(response.tokens_orig))`*%.

It's also possible to add more custom words e.g. *ACNUR*, if some organization
names are not desired, or *violencia*, if some words are overused and brings no
further explanation, in the results.

**5. Perform stemming**

Stemming is a process that removes the suffixes (and sometimes prefixes) of the
words and bring them to the base form. We use "Hunspell" stemmer from the
package [**hunspell**](https://cran.r-project.org/package=hunspell) that
provides more precise stemming behavior.

From that point onwards, we will use stemmed words instead of the raw tokenized
words because stemmed words give us better information.

```{r}
response.tokens_stemmed <- response.tokens_stopw %>% 
  mutate(word_stem = corpus::text_tokens(word, stemmer = stem_hunspell)) %>% 
  unnest(word_stem)

question.tokens_stemmed <- question.tokens_stopw %>% 
  mutate(word_stem = corpus::text_tokens(word, stemmer = stem_hunspell)) %>% 
  unnest(word_stem)
```

After stemming, the words look like this:
```{r}
rbind(head(response.tokens_stemmed[c("word", "word_stem")], 5), "...") %>% 
  gt() %>% 
  gt_condensed_style() %>% 
  cols_align(
    align = "center",
    columns = c("word", "word_stem")
  )
```

<br>

<hr>

```{r INTERNAL-prepared-data-for-further-analysis}
response.tokens <- response.tokens_stemmed
question.tokens <- question.tokens_stemmed
```

