
# How many individuals are entering data? {-}

We count the number of people entering data by each partner in *ActivityInfo*.
This information might help us derive some conclusions about the responses.

> Some responses would probably lead over time to more standardized responses
and/or maybe less mistakes as they get to know the system/meaningful inputs vs.
a more piecemeal approach (several persons entering data depending on the month
and/or no dedicated resources leading to fewer inputs, especially on the
narrative).

However, it is still not very much known if any organization (*partner*) use a
kind of "common user account" to enter either all or most of the records.

In the next **Response quality** section, we look if there are any relationship
between word counts per partner and number of people reporting to
*ActivityInfo*.

*How many unique individuals are entering narrative data from each of the
partners?*

#### Reporting partners {-}

```{r}
narratives %>% 
  select(partnerName, reportingUsers) %>% 
  tidyr::unnest(reportingUsers) %>%
  group_by(partnerName) %>% 
  summarize(numberUniqueUsers = length(unique(reportingUsers))) %>% 
  arrange(desc(numberUniqueUsers)) %>% 
  gt() %>% 
  gt_condensed_style() %>% 
  cols_label(
    partnerName = "Reporting partner",
    numberUniqueUsers = "Users"
  ) %>% 
  tab_header(
    title = "Users per reporting partner",
    subtitle = md("The number of *unique users* entering data in ActivityInfo")
  )
```

#### Implementing partners {-}

```{r}
narratives %>% 
  select(subPartnerName, reportingUsers) %>% 
  tidyr::unnest(reportingUsers) %>%
  group_by(subPartnerName) %>% 
  summarize(numberUniqueUsers = length(unique(reportingUsers))) %>% 
  arrange(desc(numberUniqueUsers)) %>% 
  gt() %>% 
  gt_condensed_style() %>% 
  cols_label(
    subPartnerName = "Implementing partner",
    numberUniqueUsers = "Users"
  ) %>% 
  tab_header(
    title = "Users per implementing partner",
    subtitle = md("The number of *unique users* entering data in ActivityInfo")
  )
```

