
# How does the term frequency change over time in sectors? {-}

We can closely look at some words to see how the term-frequency changes over
time.

Let's look at how the most frequent terms in the sectors
**VBG** (*`r unique(narratives[narratives$formNameRecode == "VBG","formName", drop=TRUE])`*),
**Poblacion** (*`r unique(narratives[narratives$formNameRecode == "Poblacion","formName", drop=TRUE])`*)
and
**Tecnico** (*`r unique(narratives[narratives$formNameRecode == "Tecnico","formName", drop=TRUE])`*),
which are
chosen after [this analysis](#average-word-count-per-sector),
change over months.

```{r}
term_sector_over_time <- function(SELECTED.FORM) {
  tf_first_per_month <- response.tokens %>% 
    group_by(formNameRecode, Month) %>% 
    count(word_stem) %>%
    dplyr::filter(formNameRecode == {{SELECTED.FORM}}) %>% 
    tidytext::bind_tf_idf(word_stem, Month, n) %>% 
    ## filter out some words that can be redundant
    dplyr::filter(!word_stem %in% c("vbg", "unfpa", "hias", "personar", "mujer")) %>% 
    ungroup() %>% 
    select(Month, word_stem, tf) %>% 
    arrange(desc(tf)) %>%
    group_by(Month) %>% 
    slice(seq(3)) %>% 
    ungroup()
  
  tf_first_per_month_ts <-
    tf_first_per_month %>%
    mutate(Month = yearmonth(Month)) %>% 
    tsibble::as_tsibble(index = "Month", key = "word_stem") %>% 
    tsibble::fill_gaps(tf = 0, .full = TRUE)
  
  tf_first_per_month_ts %>% 
    ggplot(aes(Month, tf, col = word_stem, group = word_stem)) +
    geom_point() +
    geom_line() +
    facet_grid(rows = vars(word_stem), scales = "free_y") +
    scale_fill_brewer(palette = 2) +
    scale_x_date(labels = scales::date_format("%Y-%m"), date_breaks = "1 month") +
    labs(
      title = paste0("Term frequency over time in '", SELECTED.FORM, "' sector")
    ) +
    xlab(NULL) +
    theme_ecuador1(border=TRUE, panel_spacing = 0.4) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1.1)) +
    theme(
      legend.position = "none", 
      legend.title = element_blank()
    )
}
```

```{r, fig.width=8, fig.height=25, fig.align='center', fig.cap='Term frequency over time in the sectors having the highest sample sizes.'}
cowplot::plot_grid(
  term_sector_over_time("VBG"),
  term_sector_over_time("Poblacion"),
  term_sector_over_time("Tecnico"),
  nrow = 3L
)
```

