
# Word count {-}
`TODO`
## Distribution of word counts {-}

```{r, fig.align='center'}
inspect_num(narratives.wc, breaks = 10) %>%
  show_plot()
```

There's a lot of variation in response word count. We need to know what are
those responses with a word count of 1. Or better to eliminate the responses
which has less than 5 words (an arbitrary number to filter).

`TODO correlation`

---

## Correlation {-}

The correlation between the word count of different fields in the
*ActivityInfo*:

```{r}
## Source:
## http://www.sthda.com/english/wiki/ggplot2-quick-correlation-matrix-heatmap-r-software-and-data-visualization

## Subset the numerical columns:
cordata <- narratives.wc %>% 
  select(.responseWordCount, .questionWordCount, .descriptionWordCount) %>% 
  dplyr::filter(.responseWordCount > 1)

## Compute the correlation matrix:
cormat <- round(cor(cordata), 2)
```

```{r, cor-table}
mmy::std_rownames(as.data.frame(cormat)) %>%
  gt() %>% 
  #gt_condensed_style() %>% 
  cols_label(
    .rowname = ""
  )
```

+ `r paste0(round(100*table(narratives.wc$.responseWordCount)[[1]]/length(narratives.wc$.responseWordCount),2),"%")`
of the words in the response field have only `1` words. And those words were
just not important (as they are 
`r narratives.wc %>% dplyr::filter(.responseWordCount == 1 & !is.na(response)) %>% select(response) %>% distinct() %>% pull()`).

```{r, fig.align='center'}
inspect_cor(narratives.wc) %>% 
  show_plot()
```

`TODO`

## The regression lines {-}

We can look at multiple continuous variables in our data.

- word count of response field: `response.wc`

- word count of question field: `question.wc`

- word count of description field: `description.wc`

Given this data, the model formula showing dependent and independent variables
can be as follows:

$ response.wc \sim question.wc $

$ response.wc \sim question.wc + description.wc $

In that sense the *word count of response* field is the dependent variable and
*word count of question* and *word count of description* fields are the
independent variables in the regression.

Thus, we expect to have more word count in question and description fields to
have a positive effect on the word count of response field.

Scatter plots help understand the characteristics of those variables.
However, we miss a general understanding that is the *trend line*.

```{r plot-linear-regression, fig.align='center'}
p1 <- description.field.wc %>% 
  ggplot(aes(x = .responseWordCount, y = .questionWordCount)) +
  geom_jitter(alpha = 0.35) +
  geom_smooth(aes(colour = "linear"), method = "lm") +
  theme_ecuador1() +
  theme(legend.position = "none")

p2 <- description.field.wc %>% 
  ggplot(aes(x = .responseWordCount, y = .descriptionWordCount)) +
  geom_jitter(alpha = 0.35) +
  geom_smooth(aes(colour = "linear"), method = "lm") +
  theme_ecuador1() +
  theme(legend.position = "none")

cowplot::plot_grid(p1, p2)
```

The gray area around the lines shows the confidence band at the *0.95* level.
Although there's a straight slope in the linear regression line, we cannot
say that the trend line is robust because the confidence band representing the
uncertainty in the estimate is wide.

## Logistic regression {-}

`TODO`
In fact, it is called the *binomial logistic regression*.
When one of the independent variables is dichotomous (having two categories), `...`
